For Induction-Induction, Induction is Enough

### 2019-11-11 paper/paper.tex
stuff

@@

Similar techique for other type theories: Streicher
\cite{streichersemantics}. Initiality project, Brunerie \cite{}. These
all use quotients. Closest: Lafont TYPES abstract 2017. Does not use
quotients.

@@

Natural numbers:
\[
\boldsymbol{\cdot}\,\ext\,N:\U\,\ext\,z : \El\,N\,\ext\,s : N \ra \El\,N
\]
Lists with elements of a given $A : \Set$ type. Here we use the $\Pim$ function
to include an external type into the signature.
\[
\boldsymbol{\cdot}\,\ext\,L:\U\,\ext\,nil : \El\,L\,\ext\,
      cons : \Pim\,A\,(\lambda \_.\, L\ra L)
\]
The inductive-inductive Con-Ty example from Section \ref{sec:intro}:
\begin{alignat*}{3}
  & \boldsymbol{\cdot} \,\ext\,Con : \U\,\ext\,Ty : Con\ra\U\,\ext\,empty : \El\,Con\,\ext\,ext : \Pi\,(\Gamma : Con)\,(Ty\oldapp\Gamma \ra Con) \\
  & \hspace{0.5em}\,\ext\, U : \Pi\,(\Gamma:Con)\,(\El\,(Ty\oldapp\Gamma))
    \,\ext\,El : \Pi\,(\Gamma : Con)(\El\,(Ty\oldapp(ext\oldapp\Gamma\oldapp U)))
\end{alignat*}

@@

However, we shall mention that the Agda formalisation for the current paper
cannot be directly plugged into the Agda code for
\cite{Kaposi:2019:CQI:3302515.3290315} (which includes most of the term model
construction). One reason is that the QIIT formalisation uses strict computation
rules for induction over signatures, while here we show propositional
computation. This mismatch can be in principle remedied by noting that both
formalisations use UIP and function extensionality, hence we can switch between
strict and propositional equalities, via the known translations between
extensional and intensional type
theories\cite{hofmann95extensional,winterhalter2019eliminating}.

>>>>

The current work builds heavily on the work of Kaposi et al.\ on quotient
inductive-inductive types\cite{Kaposi:2019:CQI:3302515.3290315}; we reuse both
QIIT syntax and semantics by restricting to IITs, and we reuse the term model
construction of QIITs as well.

Forsberg\cite{forsberg-phd} specifies inductive-inductive types and develops a
semantics based on dialgebras. He also provides a reduction of IITs to indexed
inductive types, but only constructs a weak version of elimination.

Hugunin\cite{jasper} constructs several IITs in cubical Agda from inductive
types. In this setting, the lack of UIP makes constructions significantly more
involved, and essentially involves coinductive-coinductive well-formedness
predicates defined as homotopy limits. Hugunin does not consider a generic
syntax of IITs and only works on specific examples (although the examples vary
greatly). He also does not construct full recursive-recursive eliminators.

Streicher\cite{streichersemantics} presents an interpretation of the well-formed
presyntax of a type theory into a categorical model, which is an important
ingredient in constructing an initial model, although he does not present
details on the construction of the term model or its initiality proof.

Voevodsky was interested in constructing initial models of type theories
from presyntaxes. There is a communal project \cite{

@@

Similar technique for other type theories: Streicher
\cite{streichersemantics}. Initiality project, Brunerie \cite{}. These
all use quotients. Closest: Lafont TYPES abstract 2017. Does not use
quotients.

@@

Natural numbers:
\[
\boldsymbol{\cdot}\,\ext\,N:\U\,\ext\,z : \El\,N\,\ext\,s : N \ra \El\,N
\]
Lists with elements of a given $A : \Set$ type. Here we use the $\Pim$ function
to include an external type into the signature.
\[
\boldsymbol{\cdot}\,\ext\,L:\U\,\ext\,nil : \El\,L\,\ext\,
      cons : \Pim\,A\,(\lambda \_.\, L\ra L)
\]
The inductive-inductive example from Section \ref{sec:intro}:
\begin{alignat*}{3}
  & \boldsymbol{\cdot} \,\ext\,Con : \U\,\ext\,Ty : Con\ra\U\,\ext\,empty : \El\,Con\,\ext\,ext : \Pi\,(\Gamma : Con)\,(Ty\oldapp\Gamma \ra Con) \\
  & \hspace{0.5em}\,\ext\, U : \Pi\,(\Gamma:Con)\,(\El\,(Ty\oldapp\Gamma))
    \,\ext\,El : \Pi\,(\Gamma : Con)(\El\,(Ty\oldapp(ext\oldapp\Gamma\oldapp (U\oldapp\Gamma))))
\end{alignat*}

@@

However, we shall mention that the Agda formalisation for the current paper
cannot be directly plugged into the Agda code for
\cite{Kaposi:2019:CQI:3302515.3290315} (which includes most of the term model
construction). One reason is that the QIIT formalisation uses strict computation
rules (given by Agda rewrite rules) for induction over signatures, while here we show propositional
computation. This mismatch can be in principle remedied by noting that both
formalisations use UIP and function extensionality, hence we can switch between
strict and propositional equalities, via the known translations between
extensional and intensional type
theories\cite{hofmann95extensional,winterhalter2019eliminating}.
